\chapter{Directives}

The \module{grok} module defines a set of directives that allow you to
configure and register your components. Most directives assume a
default, based on the environment of a module. (For example, a view
will be automatically associated with a model if the association can
be made unambigously.)

If no default can be assumed for a value, grok will explicitly tell
you what is missing and how you can provide a default or explicit
assignment for the value in question.


    \section{\function{grok.AutoFields} -- deduce schema fields automatically}

        \begin{funcdesc}{grok.AutoFields}{class_or_interface}
          A class level directive, which can be used inside
          \class{Form} classes to automatically deduce the form fields
          from the schema of the context \var{class_or_interface}.

          Different to most other directives,
          \function{grok.AutoFields} is used more like a function and
          less like a pure declaration.

          The following example makes use of the
          \function{grok.AutoFields} directive, in that one field is
          omitted from the form before rendering:

          \strong{Example:}

          \begin{verbatim}
import grok
from zope import interface, schema

class IMammoth(interface.Interface):
    name = schema.TextLine(title=u"Name")
    size = schema.TextLine(title=u"Size", default=u"Quite normal")

class Mammoth(grok.Model):
    interface.implements(IMammoth)

class Edit(grok.EditForm):
    grok.context(Mammoth)

    form_fields = grok.AutoFields(Mammoth).omit('size')
          \end{verbatim}

          In this example the \code{size} attribute will not show up
          in the resulting edit view.

          \begin{seealso}
            \class{grok.EditForm}, \class{grok.Fields}
          \end{seealso}

        \end{funcdesc}

    \section{\function{grok.adapts} -- declare that a class adapts
      certain objects}

        \begin{funcdesc}{grok.adapts}{*classes_or_interfaces}
          A class-level directive to declare that a class adapts
          objects of the classes or interfaces given in
          \var{*classes_or_interfaces}.

          This directive accepts several arguments.

          It works much like the \module{zope.component}s
          \function{adapts()}, but you do not have to make a ZCML
          entry to register the adapter.

          \strong{Example:}

          \begin{verbatim}
import grok
from zope import interface, schema
from zope.size.interfaces import ISized

class IMammoth(interface.Interface):
    name = schema.TextLine(title=u"Name")
    size = schema.TextLine(title=u"Size", default=u"Quite normal")

class Mammoth(grok.Model):
    interface.implements(IMammoth)

class MammothSize(object):
    grok.implements(ISized)
    grok.adapts(IMammoth)

    def __init__(self, context):
        self.context = context

    def sizeForSorting(self):
        return ('byte', 1000)

    def sizeForDisplay(self):
        return ('1000 bytes')
          \end{verbatim}

          Having \class{MammothSize} available, you can register it as
          an adapter, without a single line of ZCML:

          \begin{verbatim}
>>> manfred = Mammoth()
>>> from zope.component import provideAdapter
>>> provideAdapter(MammothSize)
>>> from zope.size.interfaces import ISized
>>> size = ISized(manfred)
>>> size.sizeForDisplay()
'1000 bytes'
          \end{verbatim}

          \begin{seealso}
            \class{grok.implements}
          \end{seealso}

        \end{funcdesc}

    \section{\function{grok.baseclass} -- declare a class as base}

        \begin{funcdesc}{grok.baseclass}{}
          A class-level directive without argument to mark something
          as a base class. Base classes are are not grokked.

          Another way to indicate that something is a base class, is
          by postfixing the classname with \code{'Base'}.

          The baseclass mark is not inherited by subclasses, so those
          subclasses will be grokked (except they are explicitly
          declared as baseclasses as well).

          \strong{Example:}

          \begin{verbatim}
import grok

class ModelBase(grok.Model):
    pass

class ViewBase(grok.View):
    def render(self):
        return "hello world"

class AnotherView(grok.View):
    grok.baseclass()

    def render(self):
        return "hello world"

class WorkingView(grok.View):
    pass
          \end{verbatim}

          Using this example, only the \class{WorkingView} will serve
          as a view, while calling the \class{ViewBase} or
          \class{AnotherView} will lead to a
          \exception{ComponentLookupError}.


        \end{funcdesc}

    \section{\function{grok.define_permission} -- define a permission}

        \begin{funcdesc}{grok.define_permission}{name}

          A module-level directive to define a permission with name
          \var{name}. Usually permission names are prefixed by a
          component- or application name and a dot to keep them
          unique.

          Because in Grok by default everything is accessible by
          everybody, it is important to define permissions, which
          restrict access to certain principals or roles.

          \strong{Example:}

          \begin{verbatim}
import grok
grok.define_permission('cave.enter')
          \end{verbatim}

          \begin{seealso}
            \function{grok.require()}, \class{grok.Permission},
            \class{grok.Role}
          \end{seealso}

          \versionchanged[replaced by \class{grok.Permission}]{0.11}
          
        \end{funcdesc}



    \section{\function{grok.Fields}}

        \begin{funcdesc}{grok.Fields}{*arg}
        foobar
        \end{funcdesc}

    \section{\function{grok.implements}}

        \begin{funcdesc}{grok.implements}{*arg}
        foobar
        \end{funcdesc}

    \section{\function{grok.context}}

        \begin{funcdesc}{grok.context}{*arg}
        foobar
        \end{funcdesc}

    \section{\function{grok.global_utility}}

        \begin{funcdesc}{grok.global_utility}{*arg}
        foobar
        \end{funcdesc}

    \section{\function{grok.name}}

        \begin{funcdesc}{grok.name}{*arg}
        foobar
        \end{funcdesc}

        Used to associate a component with a name. Typically this directive is
        optional. The default behaviour when no name is given depends on the
        component.

    \section{\function{grok.local_utility}}

        \begin{funcdesc}{grok.local_utility}{*arg}
        foobar
        \end{funcdesc}

    \section{\function{grok.provides}}

        \begin{funcdesc}{grok.provides}{*arg}
        foobar
        \end{funcdesc}

    \section{\function{grok.resourcedir --- XXX Not implemented yet}}

        \begin{funcdesc}{grok.resourcedir}{*arg}
        foobar
        \end{funcdesc}

        Resource directories are used to embed static resources like HTML-,
        JavaScript-, CSS- and other files in your application.

        XXX insert directive description here (first: define the name, second:
        describe the default behaviour if the directive isn't given)

        A resource directory is created when a package contains a directory
        with the name \file{static}. All files from this directory become
        accessible from a browser under the URL
        \file{http://<servername>/++resource++<packagename>/<filename>}.

        \begin{bf}Example:\end{bf} The package \module{a.b.c} is grokked and
        contains a directory \file{static} which contains the file
        \file{example.css}. The stylesheet will be available via
        \file{http://<servername>/++resource++a.b.c/example.css}.

        \begin{notice}
        A package can never have both a \file{static} directory and a Python
        module with the name \file{static.py} at the same time. grok will
        remind you of this conflict when grokking a package by displaying an
        error message.
        \end{notice}

        \subsection{Linking to resources from templates}

            grok provides a convenient way to calculate the URLs to static
            resource using the keyword \keyword{static} in page templates:

            \begin{verbatim}
<link rel="stylesheet" tal:attributes="href static/example.css" type="text/css">
            \end{verbatim}

            The keyword \keyword{static} will be replaced by the reference to
            the resource directory for the package in which the template was
            registered.

    \section{\function{grok.require}}

        \begin{funcdesc}{grok.require}{*arg}
        foobar
        \end{funcdesc}

    \section{\function{grok.site}}

        \begin{funcdesc}{grok.site}{*arg}
        foobar
        \end{funcdesc}

    \section{\function{grok.template}}

        \begin{funcdesc}{grok.template}{*arg}
        foobar
        \end{funcdesc}

    \section{\function{grok.templatedir}}

        \begin{funcdesc}{grok.templatedir}{*arg}
        foobar
        \end{funcdesc}
